---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

Author: Marie Vibeke Vestergaard

This file is defining the cohort of the study.
It also generates the outcomes to be investigated in the project and outputs phenotype and covariate file for GWAS.

Read in packages:
```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(tidyr)
library(ggpubr)
```

Paths:
```{r}
path_SSI <- "/ngc/projects2/predict_r/research/data/ssi_genotype/2023-02"
path_project <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD"
path_basefiles <- "/ngc/projects2/predict_r/research/data/basefiles/2023-07" 
path_registers <- "/ngc/projects2/predict_r/research/data/sds_registers/2022-09" 
```

## Number of all genotyped UC patients (before QC):
```{r}
#All genotyped
all <- read.csv("/ngc/projects2/predict_r/research/data/ssi_genotype/2023-02/IBD_OCT2022.fam", sep=" ", header=F)

#Load file with information on IBD patients (diagnosis and date of diagnosis)
full_ibd_cohort <- read.csv(paste0(path_project, "/GeneratedData/cohort_information/complementary_files/All_IBD_patients.csv"), sep=";")

#Combine:
colnames(all)[2] <- "lbnr"
full_ibd_cohort$lbnr <- as.character(full_ibd_cohort$lbnr)
all <- left_join(all, full_ibd_cohort, by="lbnr")

#Load file to map patients to NorDIBD versus NBS cohort:
nordibd <- read.csv(paste0(path_SSI, "/gastro_exclude.fam"), sep=" ", header=F)
all <- all %>% mutate(cohort = ifelse(lbnr %in% as.character(nordibd$V2), "NorDIBD", "NBS"))

```

## Defining the cohort:
Read in data on genotyped individuals
```{r}
#Lbnr of all genotyped individuals passing QC
df <- read.csv(paste0(path_project, "/GeneratedData/postImp/Defined_cohort_info.txt"), sep=" ", header=T) 
df <- left_join(df, all, by="lbnr")
```

Remove NBS cohort samples related to NorDIBD cohort samples:
```{r}
#Read in generated "relatedness" file:
rel <- read.csv(paste0(path_project, "/GeneratedData/QC1/related_across_cohorts_PKU.txt"), sep="\t", header=F)

df <- df %>% filter(!(lbnr %in% rel$V1))
```

Combine with full ibd cohort file to gain insight on disease:
```{r}
#Keep only UC and UC_unc (UC_unc = UC patients, but with some CD records)
df_sub <- df %>% filter(final_diag=="UC"|final_diag == "UC_unc")

#Keep only those diagnosed after 1996-01-01:
df_sub1 <- df_sub %>% filter(as.Date(diag_date)>=as.Date("1996-01-01"))

```

"End of study" date is 2022-09-01.
Remove patients without 3 years of follow-up:
```{r}
df_sub2 <- df_sub1 %>% filter(diag_date <= as.Date("2022-09-01")-years(3))
cohort <- df_sub2 %>% select(lbnr, cohort, sex, diag_age, diag_date, final_diag)

#Merge with CPR register - did any die, disappear or emigrate within the follow-up time?
cpr <- read.csv(paste0(path_registers, "/t_person.csv"), sep=";") %>% select(lbnr, C_STATUS, D_STATUS_HEN_START)
cpr$lbnr <- as.character(cpr$lbnr)
cpr <- cpr %>% filter(C_STATUS %in% c(70,80,90))
sum(duplicated(cpr$lbnr)) #each lbnr only appears once
cohort <- left_join(cohort, cpr, by = "lbnr")
cohort$diag_date <- as.Date(cohort$diag_date)  
cohort <- cohort %>% filter(!(!is.na(C_STATUS) & as.Date(D_STATUS_HEN_START)<=diag_date+years(3)))

rm(df, df_sub1, df_sub2, cpr)
```


## Generation of outcome:

Generate outcomes for the follow-up period:
Severe vs. less severe patients. Severe prognosis:
   >=5000 mg of systemic corticosteorids or
   >=2 CD-related hospitalizations (inpatient visits) > 2 days or
   >=1 CD-related major surgeries 
  
Find number of hospitalizations:
```{r}
#Read in IBD-related hospital contacts:
ibd_contacts <- read.csv(paste0(path_project, "/GeneratedData/cohort_information/complementary_files/IBD_contacts.csv"), sep=";")
ibd_contacts$D_INDDTO <- as.Date(ibd_contacts$D_INDDTO)
ibd_contacts$D_UDDTO <- as.Date(ibd_contacts$D_UDDTO)

#We are only interested in inpatient contacts longer than 2 days:
ibd_contacts <- ibd_contacts %>% filter(C_PATTYPE3 == "Inpatient") %>% filter(as.numeric(D_UDDTO-D_INDDTO)>2) %>% filter(lbnr %in% !!cohort$lbnr)
ibd_contacts$lbnr <- as.character(ibd_contacts$lbnr)

#Merge with dataset and keep only contacts within follow-up time:
ibd_contacts <- left_join(ibd_contacts, cohort, by = "lbnr")
ibd_contacts$diag_date <- as.Date(ibd_contacts$diag_date)
ibd_contacts1 <- ibd_contacts %>% filter(D_INDDTO>=diag_date & D_INDDTO<=diag_date+years(3))

#Count number of hospitalizations in period and merge with cohort
ibd_contacts3 <- ibd_contacts1 %>% group_by(lbnr) %>% summarise(outcome1a = n())
cohort <- left_join(cohort, ibd_contacts3, by = "lbnr")
cohort$outcome1a <- ifelse(is.na(cohort$outcome1a), 0, cohort$outcome1a)

```

Find number of major surgeries:
```{r}
#Read in surgery data (both LPR2 and LPR3):
ibd_surgeries <- read.csv(paste0(path_basefiles, "/ibd_patients_all_procedures.csv"), sep=";")
ibd_surgeries_lpr3 <- read.csv(paste0(path_basefiles, "/ibd_patients_all_procedures_lpr3.csv"), sep=";")

#Major surgeries (selected by Kristine Allin):
ibd_surgeries <- ibd_surgeries %>% filter(grepl(paste("KJGB", "KJFB", "KJFC", "KJFF", "KJFH", "KJFA60", "KJFA61", "KJFA63", "KJFA96", "KJFA97", "KJFW", "KJFL", sep="|"), C_OPR)) %>% filter(!(C_OPR %in% c("KJFB10", "KJFB13")))

ibd_surgeries_lpr3 <- ibd_surgeries_lpr3 %>% filter(grepl(paste("KJGB", "KJFB", "KJFC", "KJFF", "KJFH", "KJFA60", "KJFA61", "KJFA63", "KJFA96", "KJFA97", "KJFW", "KJFL", sep="|"), procedurekode)) %>% filter(!(procedurekode %in% c("KJFB10", "KJFB13")))

#Combine:
ibd_surgeries_lpr3 <- ibd_surgeries_lpr3 %>% select(lbnr, dato_start, procedurekode)
colnames(ibd_surgeries_lpr3) <- colnames(ibd_surgeries)
ibd_surgeries_all <- rbind(ibd_surgeries, ibd_surgeries_lpr3)

#Combine with cohort and keep only contacts within follow-up time:
ibd_surgeries_all$lbnr <- as.character(ibd_surgeries_all$lbnr)
ibd_surgeries_all <- left_join(ibd_surgeries_all %>% filter(lbnr %in% cohort$lbnr), cohort, by = "lbnr")
ibd_surgeries_all$D_ODTO <- as.Date(ibd_surgeries_all$D_ODTO)
ibd_surgeries_all$diag_date <- as.Date(ibd_surgeries_all$diag_date)
ibd_surgeries_all1 <- ibd_surgeries_all %>% filter(D_ODTO>=diag_date & D_ODTO<=diag_date+years(3))

#Count number of surgeries in period and merge with cohort
ibd_surgeries_all3 <- ibd_surgeries_all1 %>% group_by(lbnr) %>% summarise(outcome1b = n())
cohort <- left_join(cohort, ibd_surgeries_all3, by = "lbnr")
cohort$outcome1b <- ifelse(is.na(cohort$outcome1b), 0, cohort$outcome1b)
```


Find medication:
```{r}
medication <- read.csv(paste0(path_basefiles, "/ibd_patients_all_medications.csv"), sep=";")

#Keep only systemic corticosteroids:
medication <- medication %>% filter(atc %in% c("H02AB01", "H02AB02", "H02AB04", "H02AB06", "H02AB07", "H02AB08", "H02AB09")) %>% filter(lbnr %in% cohort$lbnr)
medication$packsize <- as.numeric(medication$packsize)
medication$strnum <- as.numeric(medication$strnum)

#Remove strunit = .
medication <- medication %>% filter(strunit != ".")

#Calculate total mg (MG and MGM (mg/ml) units for tablets and injections. The calculation is the same since "packsize" for tablets is number of tablets/package and "packsize" for inj. are number of ml)
medication <- medication %>% mutate(dosis = apk*packsize*strnum) # number of packages * tablets/ml pr package * strength of tables/ml

#Combine with cohort and keep only contacts within follow-up time:
medication$lbnr <- as.character(medication$lbnr)
medication <- left_join(medication %>% filter(lbnr %in% cohort$lbnr), cohort, by = "lbnr")
medication$eksd <- as.Date(medication$eksd)
medication$diag_date <- as.Date(medication$diag_date)
medication <- medication %>% filter(eksd>=diag_date & eksd<=diag_date+years(3))

#Count total amount of medicine in period and merge with cohort
medication <- medication %>% group_by(lbnr) %>% summarise(outcome1c = sum(dosis))
cohort <- left_join(cohort, medication, by = "lbnr")
cohort$outcome1c <- ifelse(is.na(cohort$outcome1c), 0, cohort$outcome1c)

rm(medication)
```

Combine and make binary outcome:
```{r}
cohort <- cohort %>% mutate(outcome2 = ifelse(outcome1c >= 5000 | outcome1a >= 2 | outcome1b>=1, 2, 1))

table(cohort$outcome2)

#Save file with all elements:
write.table(cohort, file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Phenotype_file_v2.csv"), sep=" ", col.names=T, quote=F, row.names=F)
```


## Make files for GWAS:

Divide into PKU and NorDIBD cohort and make file with covariates and phenotypes:
```{r}
cohort <- read.csv(file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Phenotype_file_v2.csv"), sep=" ", header=T)
cohort$lbnr <- as.character(cohort$lbnr)
cohort <- cohort%>%rename(FID=V1, IID = lbnr) %>% mutate(IIDnew = paste0(FID, "_", IID))
cohort <- cohort %>% mutate(year = year(diag_date))

#NBS cohort
cohort_NBS <- cohort %>% filter(cohort == "NBS") %>% select(FID, IID, IIDnew, sex, diag_age, year, outcome2)

#Merge with IDs and order: FID, IID, covariates, phenotypes
write.table(cohort_NBS %>% select(FID, IID, sex, diag_age, year, outcome2), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/PKU/Phenotype_file_NBS_wHeader.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#Now with the new ID
write.table(cohort_NBS %>% select(IIDnew, sex, diag_age, year,outcome2), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/PKU/Phenotype_file_NBS_wHeader_geneID.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#NorDIBD cohort
cohort_nordibd <- cohort %>% filter(cohort == "NorDIBD")%>% select(FID, IID, IIDnew, sex, diag_age,year, outcome2)

#Merge with PCs and order: FID, IID, covariates, phenotypes
write.table(cohort_nordibd %>% select(FID, IID, sex, diag_age,year, outcome2), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/NorDIBD/Phenotype_file_NorDIBD_wHeader.csv"), sep=" ", col.names=T, quote=F, row.names=F)
write.table(cohort_nordibd %>% select(IIDnew, sex, diag_age,year, outcome2), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/NorDIBD/Phenotype_file_NorDIBD_wHeader_geneID.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#Save file with FID and IID for selecting the cohort:
cohort <- rbind(cohort_NBS, cohort_nordibd)
write.table(cohort %>% select(IIDnew), paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_defined_geneID.txt"), col.names = T, row.names = F, quote = F)
write.table(cohort %>% select(FID, IID), paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_defined.txt"), col.names = T, row.names = F, quote = F)
```

## Characterise study population
```{r}
cohort <- read.csv(file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Phenotype_file_v2.csv"), sep=" ")
cohort_NBS <-  cohort %>% filter(cohort=="NBS")
cohort_nordibd <- cohort %>% filter(cohort == "NorDIBD")

#Sex
sum(cohort$sex == "F")
sum(cohort$sex == "F")/dim(cohort)[1]*100
sum(cohort$sex == "M")
sum(cohort$sex == "M")/dim(cohort)[1]*100

sum(cohort_NBS$sex == "F")
sum(cohort_NBS$sex == "F")/dim(cohort_NBS)[1]*100
sum(cohort_NBS$sex == "M")
sum(cohort_NBS$sex == "M")/dim(cohort_NBS)[1]*100

sum(cohort_nordibd$sex == "F")
sum(cohort_nordibd$sex == "F")/dim(cohort_nordibd)[1]*100
sum(cohort_nordibd$sex == "M")
sum(cohort_nordibd$sex == "M")/dim(cohort_nordibd)[1]*100

#Age at UC diagnosis
mean(cohort$diag_age)
sd(cohort$diag_age)
sum(cohort$diag_age<17)
sum(cohort$diag_age>=17 & cohort$diag_age<40)
sum(cohort$diag_age>=40)
sum(cohort$diag_age<17)/dim(cohort)[1]*100
sum(cohort$diag_age>=17 & cohort$diag_age<40)/dim(cohort)[1]*100
sum(cohort$diag_age>=40)/dim(cohort)[1]*100

sum(cohort_NBS$diag_age<17)
sum(cohort_NBS$diag_age>=17 & cohort_NBS$diag_age<40)
sum(cohort_NBS$diag_age>=40)
sum(cohort_NBS$diag_age<17)/dim(cohort_NBS)[1]*100
sum(cohort_NBS$diag_age>=17 & cohort_NBS$diag_age<40)/dim(cohort_NBS)[1]*100
sum(cohort_NBS$diag_age>=40)/dim(cohort_NBS)[1]*100

sum(cohort_nordibd$diag_age<17)
sum(cohort_nordibd$diag_age>=17 & cohort_nordibd$diag_age<40)
sum(cohort_nordibd$diag_age>=40)
sum(cohort_nordibd$diag_age<17)/dim(cohort_nordibd)[1]*100
sum(cohort_nordibd$diag_age>=17 & cohort_nordibd$diag_age<40)/dim(cohort_nordibd)[1]*100
sum(cohort_nordibd$diag_age>=40)/dim(cohort_nordibd)[1]*100

#Calendar year of diagnosis
cohort_NBS <- cohort %>% filter(cohort == "PKU")
cohort_NorDIBD <- cohort %>% filter(cohort == "NorDIBD")
sum(cohort$diag_date<as.Date("2005-01-01"))
sum(cohort$diag_date>=as.Date("2005-01-01") & cohort$diag_date<as.Date("2015-01-01"))
sum(cohort$diag_date>=as.Date("2015-01-01"))
sum(cohort$diag_date<as.Date("2005-01-01"))/dim(cohort)[1]*100
sum(cohort$diag_date>=as.Date("2005-01-01") & cohort$diag_date<as.Date("2015-01-01"))/dim(cohort)[1]*100
sum(cohort$diag_date>=as.Date("2015-01-01"))/dim(cohort)[1]*100

sum(cohort_NBS$diag_date<as.Date("2005-01-01"))
sum(cohort_NBS$diag_date>=as.Date("2005-01-01") & cohort_NBS$diag_date<as.Date("2015-01-01"))
sum(cohort_NBS$diag_date>=as.Date("2015-01-01"))
sum(cohort_NBS$diag_date<as.Date("2005-01-01"))/dim(cohort_NBS)[1]*100
sum(cohort_NBS$diag_date>=as.Date("2005-01-01") & cohort_NBS$diag_date<as.Date("2015-01-01"))/dim(cohort_NBS)[1]*100
sum(cohort_NBS$diag_date>=as.Date("2015-01-01"))/dim(cohort_NBS)[1]*100

sum(cohort_NorDIBD$diag_date<as.Date("2005-01-01"))
sum(cohort_NorDIBD$diag_date>=as.Date("2005-01-01") & cohort_NorDIBD$diag_date<as.Date("2015-01-01"))
sum(cohort_NorDIBD$diag_date>=as.Date("2015-01-01"))
sum(cohort_NorDIBD$diag_date<as.Date("2005-01-01"))/dim(cohort_nordibd)[1]*100
sum(cohort_NorDIBD$diag_date>=as.Date("2005-01-01") & cohort_NorDIBD$diag_date<as.Date("2015-01-01"))/dim(cohort_nordibd)[1]*100
sum(cohort_NorDIBD$diag_date>=as.Date("2015-01-01"))/dim(cohort_nordibd)[1]*100
```

