---
title: "Untitled"
output: html_document
date: "2023-07-28"
editor_options: 
  chunk_output_type: inline
---

Author: Marie Vibeke Vestergaard

This file runs the GWAS analyses and meta-analysis.

## Load packages and pathways in R:
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)

path_project <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD"
path_UC <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
```

## Select the cohort from the genetics data
Select the cohort:
```{bash, engine.opts='-l'}
DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData
root=$DIR/postImp
module load plink2/2.00alpha20221024
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

##Select cohort only
plink2 --vcf $root/IBDimp.1to23_related1.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --maf 0.01 --make-pgen --out $DIR/tryoutUC

#Split FID_IID: This could not be done in the original conversion of the vcf file due to the FID_gastro_control_IID samples
plink2 --pfile $DIR/tryoutUC --recode vcf --out $DIR/tryoutUC
plink2 --vcf $DIR/tryoutUC.vcf --make-pgen --id-delim "_" --out $DIR/tryoutUC

```


Read in meta data of NBS and NorDIBD cohort and make files with covariates and phenotypes:
```{r}
#NBS cohort
cohort_NBS <- read.csv(file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/PKU/Phenotype_file_NBS_wHeader.csv"), sep=" ") %>% mutate(IIDnew = paste0(FID, "_", IID)) %>% select(FID,IID, IIDnew, sex, diag_age,year, outcome2)
cohort_NBS <- cohort_NBS %>% mutate(outcome2 = ifelse(outcome2 == 2, 1, 0))

#Save files for analysis
write.table(cohort_NBS %>% select(FID,IID, sex, diag_age,year), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/PKU/Analysis_file_NBS_cov.csv"), sep=" ", col.names=T, quote=F, row.names=F)
write.table(cohort_NBS %>% select(FID, IID, outcome2), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/PKU/Analysis_file_NBS_phe.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#Make list of individuals
write.table(cohort_NBS%>% select(FID, IID), paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS.txt"), col.names = FALSE, row.names = FALSE, quote = F)
write.table(cohort_NBS%>% select(IIDnew), paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS_geneID.txt"), col.names = FALSE, row.names = FALSE, quote = F)

#NorDIBD cohort
cohort_nordibd <- read.csv(file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/NorDIBD/Phenotype_file_NorDIBD_wHeader.csv"), sep=" ") %>% mutate(IIDnew = paste0(FID, "_", IID)) %>% select(FID,IID, IIDnew, sex, diag_age,year, outcome2)

cohort_nordibd <- cohort_nordibd %>% mutate(outcome2 = ifelse(outcome2 == 2, 1, 0))

#Save files for analysis
write.table(cohort_nordibd %>% select(FID,IID, sex, diag_age,year), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/NorDIBD/Analysis_file_NorDIBD_cov.csv"), sep=" ", col.names=T, quote=F, row.names=F)
write.table(cohort_nordibd %>% select(FID,IID, outcome2), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/NorDIBD/Analysis_file_NorDIBD_phe.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#Make list of individuals
write.table(cohort_nordibd%>% select(FID,IID), paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd.txt"), col.names = FALSE, row.names = FALSE, quote = F)
write.table(cohort_nordibd%>% select(IIDnew), paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd_geneID.txt"), col.names = FALSE, row.names = FALSE, quote = F)

```

Some of variants are too sparse for NorDIBD cohort
```{bash, engine.opts='-l'}
DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData
root=$DIR/postImp

module load plink2/2.00alpha20221024
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

##Select cohort only
plink2 --pfile $DIR/tryoutUC --keep $DIR/cohort_information/UC_prognosis/Cohort_nordibd.txt --mac 5 --write-snplist --out $root/snps_nordibdUC

```

Run GWAS using regenie:
Step 1 - use data before imputation, prepared for Regenie step 1:
```{bash, engine.opts='-l'}

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/postImp
root=$DIR/IBD
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output

pheno1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/PKU
pheno2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/NorDIBD

keep1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS.txt
keep2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd.txt


module load regenie/4.3
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

regenie \
  --step 1 \
  --pgen $OUT/Regenie_step1_prep \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --out $OUT/UC_prognosis/Regenie_step1_NBS
  
#regenie \
#  --step 1 \
#  --pgen $OUT/Regenie_step1_prep \
#  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
#  --catCovarList sex \
#  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
#  --keep $keep2 \
#  --bsize 1000 \
#  --extract $DIR/snps_nordibdUC.snplist \
#  --bt \
#  --out $OUT/UC_prognosis/Regenie_step1_NorDIBD


```

Step 2:
```{bash, engine.opts='-l'}

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/postImp
root=$DIR/IBD
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output

pheno1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/PKU
pheno2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/NorDIBD

keep1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS.txt
keep2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd.txt


module load regenie/4.3
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

regenie \
  --step 2 \
  --pgen $DIR/tryoutUC \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/UC_prognosis/Regenie_step1_NBS_pred.list \
  --out $OUT/UC_prognosis/Regenie_step2_NBS
  
#regenie \
#  --step 2 \
#  --pgen $DIR/tryoutUC \
#  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
#  --catCovarList sex \
#  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
#  --keep $keep2 \
#  --bsize 1000 \
#  --bt \
#  --firth \
#  --extract $DIR/snps_nordibdUC.snplist \
#  --pThresh 0.01 \
#  --pred $OUT/UC_prognosis/Regenie_step1_NorDIBD_pred.list \
#  --out $OUT/UC_prognosis/Regenie_step2_NorDIBD


```

Prepare output for meta-analysis:
Required: CHR, BP, SNP, A1, A2, P, OR, SE
```{r}
path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"

gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)

write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
#rm(gwas_NBS)

#NorDIBD
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)

write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
#rm(gwas_nord)

```


Run meta-analyses using plink1.9:
```{bash, engine.opts='-l'}

OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis


module load plink2/1.90beta6.18
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

###############
#Meta-analyses#
###############

plink --meta-analysis $OUT/Regenie_step2_NBS_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_outcome2_prep.regenie + logscale --out $OUT/meta_regenie
```


## Investigate and plot results in R

Read in libraries and pathways
```{r}
library(dplyr)
library(ggplot2)
library(qqman)
library(ggpubr)

path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
```

Calculate inflation factors, find genome-wide significant results and top 10 SNPs with lowest p-values:
```{r}
meta <- read.table(file = paste0(path, "/meta_regenie.meta"), header=T)
meta$SE <- abs(log(meta$OR)/qnorm(meta$P/2))

#Calculate inflation factor
inf <-  median(qchisq(meta$P, df=1, lower.tail = F))/0.4549

#Find significant for table
meta_sig <-meta %>% filter(P<5*10^-8)
```

## Run GWAS on HLA data:
Prepare HLA data:
```{bash, engine.opts='-l'}
DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData
root=$DIR/postImp

HLA=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/hla_imp
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis

module load plink2/2.00alpha20221024
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

#DRB1
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_DRB1.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_DRB1

plink2 --pfile $OUT/UC_HLA_DRB1 --recode vcf --out $OUT/UC_HLA_DRB1
plink2 --vcf $OUT/UC_HLA_DRB1.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_DRB1

#DPB1
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_DPB1.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_DPB1

plink2 --pfile $OUT/UC_HLA_DPB1 --recode vcf --out $OUT/UC_HLA_DPB1
plink2 --vcf $OUT/UC_HLA_DPB1.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_DPB1

#DQB1
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_DQB1.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_DQB1

plink2 --pfile $OUT/UC_HLA_DQB1 --recode vcf --out $OUT/UC_HLA_DQB1
plink2 --vcf $OUT/UC_HLA_DQB1.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_DQB1

#C
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_C.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_C

plink2 --pfile $OUT/UC_HLA_C --recode vcf --out $OUT/UC_HLA_C
plink2 --vcf $OUT/UC_HLA_C.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_C

#B
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_B.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_B

plink2 --pfile $OUT/UC_HLA_B --recode vcf --out $OUT/UC_HLA_B
plink2 --vcf $OUT/UC_HLA_B.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_B

#A
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_A.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_A

plink2 --pfile $OUT/UC_HLA_A --recode vcf --out $OUT/UC_HLA_A
plink2 --vcf $OUT/UC_HLA_A.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_A

#DQA1
plink2 --vcf $HLA/IBDimp_X_unphased_related.hla_DQA1.vcf.gz --keep $DIR/cohort_information/UC_prognosis/Cohort_defined_geneID.txt --make-pgen --out $OUT/UC_HLA_DQA1

plink2 --pfile $OUT/UC_HLA_DQA1 --recode vcf --out $OUT/UC_HLA_DQA1
plink2 --vcf $OUT/UC_HLA_DQA1.vcf --make-pgen --id-delim "_" --out $OUT/UC_HLA_DQA1

```

Regenie Step 2:
```{bash, engine.opts='-l'}

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/postImp
root=$DIR/IBD
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis

pheno1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/PKU
pheno2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/NorDIBD

keep1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS.txt
keep2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd.txt

HLA=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/hla_imp

module load regenie/4.3
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

#DRB1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DRB1
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DRB1

#DPB1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DPB1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DPB1
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DPB1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DPB1

#DQB1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQB1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DQB1
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQB1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DQB1

#DQA1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQA1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DQA1
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQA1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DQA1

#C
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_C \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_C
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_C \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_C

#B
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_B \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_B
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_B \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_B

#A
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_A \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_A
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_A \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_A

```

Prepare output for meta-analysis:
Required: CHR, BP, SNP, A1, A2, P, OR, SE
```{r}
path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
#DRB1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DRB1_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DRB1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DRB1_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DRB1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#DPB1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DPB1_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DPB1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DPB1_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DPB1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#DQB1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DQB1_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DQB1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DQB1_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DQB1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#DQA1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DQA1_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DQA1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DQA1_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DQA1_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#C
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_C_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_C_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_C_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_C_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#B
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_B_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_B_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_B_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_B_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#A
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_A_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_A_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_A_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_A_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

```


Run meta-analyses using plink1.9:
```{bash, engine.opts='-l'}

OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis


module load plink2/1.90beta6.18
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

###############
#Meta-analyses#
###############

#DRB1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DRB1_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DRB1_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DRB1

#DPB1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DPB1_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DPB1_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DPB1

#DQB1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DQB1_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DQB1_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DQB1

#DQA1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DQA1_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DQA1_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DQA1

#C
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_C_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_C_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_C

#B
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_B_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_B_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_B

#A
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_A_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_A_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_A
```


## Investigate and plot results in R

Read in libraries and pathways
```{r}
library(dplyr)
library(ggplot2)
library(qqman)
library(ggpubr)

path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
```

Find genome-wide significant results:
```{r}
hla_imp <- c("DRB1", "DPB1", "DQB1", "DQA1", "C", "B", "A")
meta <- data.frame()
for (hla in hla_imp){
  met <- read.table(file = paste0(path, "/meta_regenie_HLA_", hla, ".meta"), header=T)
  meta <- rbind(meta, met)
}

meta_sig <- meta %>% filter(P<5*10^-8)

```

## GWAS conditioning on HLA-DRB1*01:03:
Take out genotype of DRB1*01:03:
```{bash, engine.opts='-l'}
DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData
root=$DIR/postImp

HLA=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/hla_imp
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis

module load plink2/2.00alpha20221024
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

#DRB1
plink2 --pfile $OUT/UC_HLA_DRB1 --snp HLA-DRB1*01:03 --make-pgen --out $OUT/UC_conditional

```

Step 2:
```{bash, engine.opts='-l'}

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/postImp
root=$DIR/IBD
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output

pheno1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/PKU
pheno2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/NorDIBD

keep1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS.txt
keep2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd.txt


module load regenie/4.3
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

regenie \
  --step 2 \
  --pgen $DIR/tryoutUC \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --condition-file pgen,$OUT/UC_prognosis/UC_conditional \
  --condition-list $OUT/UC_prognosis/condition_list.txt \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/UC_prognosis/Regenie_step1_NBS_pred.list \
  --out $OUT/UC_prognosis/Regenie_NBS_conditional
  
regenie \
  --step 2 \
  --pgen $DIR/tryoutUC \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --condition-file pgen,$OUT/UC_prognosis/UC_conditional \
  --condition-list $OUT/UC_prognosis/condition_list.txt \
  --bsize 1000 \
  --bt \
  --firth \
  --extract $DIR/snps_nordibdUC.snplist \
  --pThresh 0.01 \
  --pred $OUT/UC_prognosis/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/UC_prognosis/Regenie_NorDIBD_conditional


```

Prepare output for meta-analysis:
Required: CHR, BP, SNP, A1, A2, P, OR, SE
```{r}
path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"

gwas_NBS <- read.table(file = paste0(path, "/Regenie_NBS_conditional_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_NBS_conditional_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
#rm(gwas_NBS)

#NorDIBD
gwas_nord <- read.table(file = paste0(path, "/Regenie_NorDIBD_conditional_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_NorDIBD_conditional_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
#rm(gwas_nord)

```


Run meta-analyses using plink1.9:
```{bash, engine.opts='-l'}

OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis


module load plink2/1.90beta6.18
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

###############
#Meta-analyses#
###############

plink --meta-analysis $OUT/Regenie_NBS_conditional_outcome2_prep.regenie $OUT/Regenie_NorDIBD_conditional_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_conditional
```


## Investigate and plot results in R

Read in libraries and pathways
```{r}
library(dplyr)
library(ggplot2)
library(qqman)
library(ggpubr)

path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
```

Calculate inflation factors, find genome-wide significant results:
```{r}
meta <- read.table(file = paste0(path, "/meta_regenie_conditional.meta"), header=T)
meta$SE <- abs(log(meta$OR)/qnorm(meta$P/2))


#Calculate inflation factor
inf <-  median(qchisq(meta$P, df=1, lower.tail = F))/0.4549

#Find significant for table
meta_sig <-meta %>% filter(P<5*10^-8)

meta %>% filter(SNP=="rs144717024")
```

Extract DRB1*01:03 genotype for association tests:
```{bash, engine.opts='-l'}
#!/bin/bash

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/UC_analysis
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis
module load plink2/2.00alpha20221024

#Recode
plink2 --pfile $OUT/UC_HLA_DRB1 --snp HLA-DRB1*01:03 --recode A --window 0  --out $OUT/snp_HLA

```

## Conditioning analysis on the HLA region:
```{bash, engine.opts='-l'}

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/postImp
root=$DIR/IBD
OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis

pheno1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/NBS
pheno2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/NorDIBD

keep1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_NBS.txt
keep2=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis/Cohort_nordibd.txt

HLA=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/hla_imp

module load regenie/4.3
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

#DRB1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DRB1_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DRB1_cond

#DPB1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DPB1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DPB1_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DPB1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DPB1_cond

#DQB1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQB1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DQB1_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQB1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DQB1_cond

#DQA1
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQA1 \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_DQA1_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DQA1 \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_DQA1_cond

#C
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_C \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_C_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_C \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_C_cond

#B
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_B \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_B_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_B \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_B_cond

#A
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_A \
  --covarFile $pheno1/Analysis_file_NBS_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno1/Analysis_file_NBS_phe.csv \
  --keep $keep1 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NBS_pred.list \
  --out $OUT/Regenie_step2_NBS_HLA_A_cond
  
regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_A \
  --covarFile $pheno2/Analysis_file_NorDIBD_cov.csv \
  --catCovarList sex \
  --phenoFile $pheno2/Analysis_file_NorDIBD_phe.csv \
  --keep $keep2 \
  --bsize 1000 \
  --condition-file pgen,$OUT/UC_conditional \
  --condition-list $OUT/condition_list.txt \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_NorDIBD_pred.list \
  --out $OUT/Regenie_step2_NorDIBD_HLA_A_cond

```

Prepare output for meta-analysis:
Required: CHR, BP, SNP, A1, A2, P, OR, SE
```{r}
path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
#DRB1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DRB1_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DRB1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DRB1_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DRB1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#DPB1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DPB1_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DPB1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DPB1_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P))
gwas_nord <- gwas_nord %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DPB1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#DQB1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DQB1_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DQB1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DQB1_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DQB1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#DQA1
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_DQA1_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_DQA1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_DQA1_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_DQA1_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#C
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_C_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P))%>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_C_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_C_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_C_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#B
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_B_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_B_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_B_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_B_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

#A
gwas_NBS <- read.table(file = paste0(path, "/Regenie_step2_NBS_HLA_A_cond_outcome2.regenie"), header=T)
gwas_NBS <- gwas_NBS %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_NBS) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_NBS <- gwas_NBS %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_NBS,  paste0(path, "/Regenie_step2_NBS_HLA_A_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)
gwas_nord <- read.table(file = paste0(path, "/Regenie_step2_NorDIBD_HLA_A_cond_outcome2.regenie"), header=T)
gwas_nord <- gwas_nord %>% select(CHROM, GENPOS, ID, ALLELE0, ALLELE1, BETA, SE, LOG10P)
colnames(gwas_nord) <- c("CHR", "BP", "SNP", "A1", "A2", "BETA", "SE", "LOG10P")
gwas_nord <- gwas_nord %>% mutate(P = 10^(-LOG10P)) %>%select(CHR, SNP, BP, A1, A2, P, BETA, SE)
write.table(gwas_nord,  paste0(path, "/Regenie_step2_NorDIBD_HLA_A_cond_outcome2_prep.regenie"), sep=" ", col.names=T, quote=F, row.names=F)

```


Run meta-analyses using plink1.9:
```{bash, engine.opts='-l'}

OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis


module load plink2/1.90beta6.18
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

###############
#Meta-analyses#
###############

#DRB1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DRB1_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DRB1_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DRB1_cond

#DPB1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DPB1_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DPB1_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DPB1_cond

#DQB1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DQB1_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DQB1_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DQB1_cond

#DQA1
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_DQA1_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_DQA1_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_DQA1_cond

#C
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_C_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_C_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_C_cond

#B
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_B_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_B_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_B_cond

#A
plink --meta-analysis $OUT/Regenie_step2_NBS_HLA_A_cond_outcome2_prep.regenie $OUT/Regenie_step2_NorDIBD_HLA_A_cond_outcome2_prep.regenie + logscale --out $OUT/meta_regenie_HLA_A_cond
```


## Investigate and plot results in R

Read in libraries and pathways
```{r}
library(dplyr)
library(ggplot2)
library(qqman)
library(ggpubr)

path <- "/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis"
```

Calculate inflation factors, find genome-wide significant results:
```{r}
hla_imp <- c("DRB1", "DPB1", "DQB1", "DQA1", "C", "B", "A")
meta <- data.frame()
for (hla in hla_imp){
  met <- read.table(file = paste0(path, "/meta_regenie_HLA_", hla, "_cond.meta"), header=T)
  meta <- rbind(meta, met)
}
meta %>% arrange(P) %>% head()
meta_sig <- meta %>% filter(P<5*10^-8)

```

## Rerun analyses separately for the different outcome criteria for the HLA variant:

Make files:
```{r}
cohort <- read.csv(file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Phenotype_file_v2.csv"), sep=" ", header=T) 
cohort$lbnr <- as.character(cohort$lbnr)
cohort <- cohort %>% mutate(year = year(diag_date))
cohort <- cohort %>% rename(IID =lbnr)
cohort <- left_join(cohort, ids, by="IID")

#Generate binary outcomes - excluding other types of cases from the controls:
cohort <- cohort %>% mutate(hosp = ifelse(outcome1a>=2, 1, 0), surg = ifelse(outcome1b>=1, 1, 0), cortico = ifelse(outcome1c>=5000, 1, 0))

#Subset cohorts
#This removes controls, that are not controls in original analysis
cohort_hosp <- cohort %>% filter(hosp==1 | (hosp==0 & outcome2==1) )
cohort_surg <- cohort %>% filter((surg==0 & outcome2==1) | surg==1)
cohort_cortico <- cohort %>% filter((cortico==0 & outcome2==1) | cortico==1)

table(cohort_hosp$hosp)
table(cohort_surg$surg)
table(cohort_cortico$cortico)

#Save files for analysis
write.table(cohort_hosp %>% select(FID, IID), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_hosp.txt"), sep=" ", col.names=T, quote=F, row.names=F)
write.table(cohort_surg %>% select(FID, IID), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_surg.txt"), sep=" ", col.names=T, quote=F, row.names=F)
write.table(cohort_cortico %>% select(FID, IID), file = paste0(path_project, "/GeneratedData/cohort_information/UC_prognosis/Cohort_cortico.txt"), sep=" ", col.names=T, quote=F, row.names=F)
```

Step 1 and 2:
```{bash, engine.opts='-l'}

DIR=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/postImp
root=$DIR/IBD
OUT1=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output

OUT=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/Output/UC_prognosis

pheno=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis


keep=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/cohort_information/UC_prognosis

HLA=/ngc/projects2/predict_r/research/projects/0007_Genetics_SNP_QC_IMP_PKU_NordIBD/GeneratedData/hla_imp


module load regenie/4.3
module load tools
module load gcc
module load intel/perflibs
module load R/4.0.0

#Hosp
regenie \
  --step 1 \
  --pgen $OUT1/Regenie_step1_prep \
  --covarFile $pheno/Analysis_file_cov_sep_bin.csv \
  --catCovarList sex,cohort \
  --phenoFile $pheno/Analysis_file_phe_sep_bin.csv --phenoCol hosp \
  --keep $keep/Cohort_hosp.txt \
  --bsize 1000 \
  --bt \
  --out $OUT1/UC_prognosis/Regenie_step1_hosp
  

regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno/Analysis_file_cov_sep_bin.csv \
  --catCovarList sex,cohort \
  --phenoFile $pheno/Analysis_file_phe_sep_bin.csv  --phenoCol hosp \
  --keep $keep/Cohort_hosp.txt \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_hosp_pred.list \
  --out $OUT/Regenie_step2_HLA_DRB1_hosp

#Surg
regenie \
  --step 1 \
  --pgen $OUT1/Regenie_step1_prep \
  --covarFile $pheno/Analysis_file_cov_sep_bin.csv \
  --catCovarList sex,cohort \
  --phenoFile $pheno/Analysis_file_phe_sep_bin.csv --phenoCol surg \
  --keep $keep/Cohort_surg.txt \
  --bsize 1000 \
  --bt \
  --out $OUT1/UC_prognosis/Regenie_step1_surg
  

regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno/Analysis_file_cov_sep_bin.csv \
  --catCovarList sex,cohort \
  --phenoFile $pheno/Analysis_file_phe_sep_bin.csv  --phenoCol surg \
  --keep $keep/Cohort_surg.txt \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_surg_pred.list \
  --out $OUT/Regenie_step2_HLA_DRB1_surg


#cortico
regenie \
  --step 1 \
  --pgen $OUT1/Regenie_step1_prep \
  --covarFile $pheno/Analysis_file_cov_sep_bin.csv \
  --catCovarList sex,cohort \
  --phenoFile $pheno/Analysis_file_phe_sep_bin.csv --phenoCol cortico \
  --keep $keep/Cohort_cortico.txt \
  --bsize 1000 \
  --bt \
  --out $OUT1/UC_prognosis/Regenie_step1_cortico
  

regenie \
  --step 2 \
  --pgen $OUT/UC_HLA_DRB1 \
  --covarFile $pheno/Analysis_file_cov_sep_bin.csv \
  --catCovarList sex,cohort \
  --phenoFile $pheno/Analysis_file_phe_sep_bin.csv  --phenoCol cortico \
  --keep $keep/Cohort_cortico.txt \
  --bsize 1000 \
  --bt \
  --firth \
  --pThresh 0.01 \
  --pred $OUT/Regenie_step1_cortico_pred.list \
  --out $OUT/Regenie_step2_HLA_DRB1_cortico

```

